### 2023-12-21 manuscript.tex
fix references [ci skip]
This build is based on
https://github.com/MMV-Lab/im2im-paper/commit/dd4b489eea3392962ffb0d1388911c331d929bbb.

This commit was created by the following CI build and job:
https://github.com/MMV-Lab/im2im-paper/commit/dd4b489eea3392962ffb0d1388911c331d929bbb/checks
https://github.com/MMV-Lab/im2im-paper/actions/runs/7286776855

This manuscript
(\href{https://MMV-Lab.github.io/im2im-paper/v/2a877704ba5a3a36e81963a66743cc14dd0d0786/}{permalink})
was automatically generated
from \href{https://github.com/MMV-Lab/im2im-paper/tree/2a877704ba5a3a36e81963a66743cc14dd0d0786}{MMV-Lab/im2im-paper@2a87770}
on December 21, 2023.

@@

For computational biomedical researchers, \emph{MMV\_Im2Im} provides a starting point for developing new biomedical image analysis or machine learning algorithms, where they can either reuse the code in this package or fork and extend this package to facilitate the development of new methods. Experimental biomedical researchers can benefit from this work by gaining a comprehensive view of the image-to-image transformation concept through diversified examples and use cases. We hope this work can give the community inspirations on how DL-based image-to-image transformation can be integrated into the assay development process, enabling new biomedical studies that cannot be done only with traditional experimental assays. To help researchers get started, we have provided source code, documentation, and tutorials for \emph{MMV\_Im2Im} at {[}\protect\hyperlink{ref-github_mmv_im2im}{\textbf{github\_mmv\_im2im?}}{]} under MIT license.

@@

The original idea of a general toolbox actually stemmed from the OpenMMLab project (\url{https://openmmlab.com/}), which provides generic codebases for a wide range of computer vision research topics. For instance, \emph{MMSegmentation} {[}\protect\hyperlink{ref-EcizztLg}{16}{]} is an open source toolbox for semantic segmentation, supporting unified benchmarking and state-of-the-art models ready to use out-of-box. It has become one of most widely used codebase for research in semantic segmentation (2.3K forks and 6.5K stars on GitHub as of September 29, 2023). This inspires us to develop \emph{MMV\_Im2Im} to facilitate research in image-to-image transformation with a special focus on biomedical applications.

First of all, different from general computer vision datasets, such as ImageNet {[}\protect\hyperlink{ref-lt4BNUoG}{17}{]}, where the images are usually small 2D RGB images (e.g., 3 x 256 x 256 pixels), biomedical applications usually involves large-scale high dimensional data (e.g., 500 images of 4 x 128 x 2048 x 2048 voxels). To deal with this issue, we employ the PersistentDataset in MONAI {[}\protect\hyperlink{ref-UU62HYC6}{18}{]} with partial loading and sampling support, as well as delayed image reading powered by aicsimageio {[}\protect\hyperlink{ref-gsfWGJKf}{19}{]} as default (configurable if another dataloader is preferred). As a result, in our stress test, training a 3D nuclei instance segmentation model with more than 125,000 3D images can be conducted efficiently in a day, even with limited resources.

@@

Third, the toolbox pre-packs common functionalities specific to microscopy images. For example, we incorporate the special image normalization method introduced in {[}\protect\hyperlink{ref-Yq8wZ6hc}{4}{]}, where only the middle chunk along the Z dimension of 3D microscopy images will be used for calculating the mean and standard deviation of image intensity for standard normalization. Also, 3D light microscopy images are usually anisotropic, i.e., much lower resolution along Z than XY dimension. So, we adopt the anisotropic variation of UNet as proposed in {[}\protect\hyperlink{ref-jM3v1UjQ}{20}{]}.

Finally, to deploy the model in production, a model trained on small 3D patches sometimes needs to be applied not only on much larger images. Combining the efficient data handling of aicsimageio {[}\protect\hyperlink{ref-gsfWGJKf}{19}{]} and the sliding window inference with gaussian weighted blending, the toolbox can yield efficient inference without visible stitching artifacts in production.

All in all, the \emph{MMV\_Im2Im} toolbox stands on the shoulders of many giants in the open-source software and ML engineering communities (pytorch-lightning, MONAI, aicsimageio, etc.) and is systematically designed for image-to-image transformation R\&D for biomedical applications. The source code of \emph{MMV\_Im2Im} is available at {[}\protect\hyperlink{ref-github_mmv_im2im}{\textbf{github\_mmv\_im2im?}}{]}. This manuscript is generated with the open-source package Manubot {[}\protect\hyperlink{ref-YuJbg3zO}{21}{]}. The manuscript source code is available at {[}\protect\hyperlink{ref-github_manuscript}{\textbf{github\_manuscript?}}{]}.

@@

We did a special comparison in this subsection to further illustrate the difference between semantic and instance segmentations. We took the 3D fibrillarin dataset from {[}\protect\hyperlink{ref-5sGcmDuy}{26}{]}. There are multiple channels in each 3D image, including DNA dye, membrane dye, and the structure channel (i.e., fibrillarin in this case). The original fibrillarin segmentation released with the dataset is a semantic segmentation (0=background, 1=fibrillarin). With the additional cell segmentation available in the dataset, we can know which groups of segmented fibrillarin belong to the same cell. Then, we can convert the original 3D fibrillarin semantic segmentation ground truth into 3D instance segmentation ground truth (fibrillarin pixels belonging to the same cell are grouped as a unique instance). Sample images and results are shown in Figure \ref{fig:3dseg}. We can observe that the semantic segmentation model is able to achieve good accuracy in determining pixels from the fibrillarin signals (F1 = 0.958 ± 0.008). Meanwhile, the instance segmentation can group them properly (average precision at 50 = 0.795 ± 0.055) so that fibrillarin masks from the same cell are successfully identified as unique instances, even without referring to the cell membrane channel or cell segmentation results. This is not a simple grouping step based on distance, since the fibrillarin signals from tightly touching nuclei may exist very close to each other.

@@

Large amounts of high-quality segmentation ground truth is not always available, or may require endless effort to collect for a segmentation task. CycleGAN-based methods have opened up a new avenue for segmentation without the need for pixel-wise ground truth {[}\protect\hyperlink{ref-RuFP3CS3}{12}{]}. In this subsection, we demonstrate an unsupervised learning-based segmentation method on four examples: 2D tight-junction (via ZO1) segmentation from 2D FP-tagged ZO1 images (max-projected from 3D stacks), and segmentation of nuclei, mitochondria, and golgi from 3D confocal microscopy images.

To perform unsupervised learning, we used raw images from the hiPSC single-cell image dataset {[}\protect\hyperlink{ref-5sGcmDuy}{26}{]}, as well as their corresponding segmentations (may not be absolute pixel-wise ground truth, but have gone through systematic evaluation to ensure the overall quality). We shuffled the raw images and their segmentations to generate a set of simulated segmentation masks. A demonstration of the concept is illustrated in Figure \ref{fig:unsupervised}-A. Example results for all 3D models are shown in Figure \ref{fig:unsupervised}-B, and the F1-scores on the test set are summarized in Table \ref{tbl:unsuper}.

For the 2D example, we saw that the unsupervised training provides a valuable segmentation, which is reflected by the F1 Score in Table \ref{tbl:unsuper}. For the 3D examples, it has been suggested that the quality of unsupervised nuclei segmentation could be further improved with additional simulation strategies {[}\protect\hyperlink{ref-RuFP3CS3}{12}{]}. Overall, we believe that unsupervised learning offers an effective way to generate preliminary segmentation, which can be further refined through active learning such as the iterative DL workflow described in {[}\protect\hyperlink{ref-jM3v1UjQ}{20}{]}.

@@

Generating a large amount of synthetic microscopy images can be an important step in developing image analysis methods. Synthetic images offer a way to train other DL models, such as self-supervised pre-training, using a diverse set of images without the need for large amounts of real-world data. As long as the synthetic images are generated with sufficient quality, it is possible to have an unlimited amount of training data for certain applications. Moreover, synthetic images can be used to evaluate other models when validation data is difficult to obtain. In this study, we demonstrate that \emph{MMV\_Im2Im} can generate 2D/3D synthetic microscopy images with high realism and validity, using a subset of data collected from the hiPSC single-cell image dataset {[}\protect\hyperlink{ref-5sGcmDuy}{26}{]}, either in a supervised or unsupervised manner.

@@

\emph{MMV\_Im2Im} can also be used to computationally reduce image noise or restore the data from various sources of imaging artifacts, so as to increase the feasibility and efficiency in downstream analysis. In the current version of \emph{MMV\_Im2Im}, the restoration model can only be trained in a fully supervised manner. Therefore, aligned low-quality and high-quality images are required for supervision, even though such paired data can be partially simulated {[}\protect\hyperlink{ref-wcCVn8av}{6}{]}. Other methods, such as unsupervised learning-based solutions {[}\protect\hyperlink{ref-4vnyY9J9}{42}{]}, will be made available within \emph{MMV\_Im2Im} in future versions.

In this example, we presented an image denoising demonstration with sample data from {[}\protect\hyperlink{ref-12G712Zky}{43}{]}. The goal was to increase the quality of low signal-to-noise ratio (SNR) images of nucleus-stained flatworms (Schmidtea mediterranea, planaria) and lightsheet images of Tribolium castaneum (red flour beetle) embryos. The models were trained with paired data acquired with low and high laser intensity on fixed samples, and then applied on live imaging data. For the nucleus-stained flatworm data (a test set of 20 images are available), the model achieved Pearson correlation of 0.392 ± 0.065, while the Pearson correlation between the noisy raw and ground truth images was 0.065 ± 0.057. For the red flour beetle dataset, the model has improved the Pearson correlation from 0.219 ± 0.045 to 0.444 ± 0.077 (6 images). Based on this and the results in Figure \ref{fig:denoising}, it can be observed that the low SNR images can be greatly improved. Systematic quantitative evaluations would be necessary to confirm the biological validity, but beyond the scope of this paper.

@@

Another important application of image-to-image transformation is imaging modality transformation {[}\protect\hyperlink{ref-UEBDZ3tI}{8}{]}, usually from one ``cheaper'' modality with lower resolution (e.g., with larger field-of-view, easier to acquire and scale up) to another modality with higher resolution but expensive to obtain. Such models will permit a new way in assay development strategy to take advantage of all the benefits of the cheaper modality with lower resolution and still be able to enhance the resolution computationally post hoc. To demonstrate the application of \emph{MMV\_Im2Im} in this scenario, we took an example dataset with paired 3D confocal and Stimulated Emission Depletion (STED) images of two different cellular structures, microtubule and nuclear pore {[}\protect\hyperlink{ref-UEBDZ3tI}{8}{]}. Sample results were summarized in Figure \ref{fig:modality_mt} and Figure \ref{fig:modality_np}. The corresponding error plots show pixel-based absolute differences between ground truth and prediction. Intensities were normalized to the interval from -1 to 1 for training, with intensity limits restricted to the 0.01-percentile and 99.99-percentile values of the intensity distribution.

@@

DL has emerged as a powerful tool for multiplex imaging, a powerful technique that enables the simultaneous detection and visualization of multiple biomolecules within a single tissue sample. This technique is increasingly being used in biomedical experiments but demands efficient image analysis solutions to accurately identify and quantify the different biomolecules of interest at scale. DL has demonstrated great potentials in analyzing multiplex datasets, as it can automatically learn the complex relationships between different biomolecules and their spatial distribution within tissues. Specifically, in this study, we present the effectiveness of \emph{MMV\_Im2Im} in transforming tissue images from one staining to another, which will permit efficient co-registration, co-localization, and quantitative analysis of multiplex datasets. We used the sample dataset from {[}\protect\hyperlink{ref-WwenuBHa}{5}{]}. In this example, we trained three different models to transform IHC images to images of standard hematoxylin stain, mpIF nuclear (DAPI) and mpIF LAP2beta (a nuclear envelope stain). Example results can be observed in Figure \ref{fig:multiplex} to verify the results qualitatively, the respective metrics can be found in Table \ref{tbl:multiplex}. It is worth mentioning that there is a pixel shift in the mpIF LAP2beta holdout dataset, but image registration is beyond the scope of this manuscript. We show the metrics as an example of an evaluation of the transformed images, but we leave an application-specific evaluation to the appropriate researchers. But it is evident that these transformed images can provide valuable insights into the localization and expression patterns of specific biomolecules spatially.

@@

From all experiments above (37 in total), we want to demonstrate the great flexibility of \emph{MMV\_Im2Im} and not to optimize every task in detail. Presenting all detailed configurations in these 37 experiments in the manuscript could lead to more confusion than clarity. To this end, we give a high-level overview of the key information of each task in Table \ref{tbl:framework_overview}, hoping to serve as a valuable starting point for researchers to optimize their DL-based image-to-image transformation using \emph{MMV\_Im2Im}. The full configuration details are available in human-readable formats in our GitHub repository {[}\protect\hyperlink{ref-github_mmv_im2im}{\textbf{github\_mmv\_im2im?}}{]}.

@@

Overall, the package inherited the boilerplate concept from pytorch-lightning {[}\protect\hyperlink{ref-YbvSvdyB}{14}{]}, and was made fully configurable via yaml files supported by pyrallis {[}\protect\hyperlink{ref-17WhsEZko}{44}{]}, as well as largely employed state-of-the-art DL components from MONAI {[}\protect\hyperlink{ref-UU62HYC6}{18}{]}. The three key parts in the package: \texttt{mmv\_im2im.models}, \texttt{mmv\_im2im.data\_modules}, and \texttt{Trainers}, will be further described below.

@@

In general, there are mainly four major DL frameworks that could be applied to microscopy image-to-image transformation: supervised learning with a fully convolutional networks (FCN) type models, supervised learning with pix2pix type models, unsupervised learning to learn mapping between visual domains, and Self2Self-type self-supervised learning {[}\protect\hyperlink{ref-tuObtXMR}{45}{]}. The major difference between FCN based supervised learning and pix2pix based supervised learning is that the pix2pix framework extends an FCN model with an adversarial head as a discriminator to further improve the realism of the prediction. The major difference between the unsupervised framework and the self-supervised framework is that the unsupervised methods still require examples of the target images, even though the source images and target images do not need to be from the same sample or pixel-wise aligned. But, the self-supervised framework would only need the original images, which could be really helpful when it is impossible to acquire the target images (e.g., there is no truly noise-free or artifact-free image).

Currently, for supervised frameworks, both the FCN-type and pix2pix-type are well supported in the MMV\_Im2Im (RRID:SCR\_024630) package. Since our package is designed in a very generic way, it is possible to continuously expand the functionalities when available (ideally with community contributions). For example, diffusion models {[}\protect\hyperlink{ref-1A3yurr7m}{46}{]} can be thought of as a modern extension of the pix2pix-type framework and therefore are within our horizon to include into \emph{MMV\_Im2Im}. For the unsupervised framework, only CycleGAN-type methods are supported. We are planning to extend the unsupervised framework with Imaginaire {[}\protect\hyperlink{ref-vzOBQiEH}{47}{]}, which will greatly extend the applicability of \emph{MMV\_Im2Im} (e.g., learning the transformation from one single image to another single image or one set of images to another set of images). Meanwhile, supporting the self-supervised framework will be our next major milestone.

@@

One of main directions for extending \emph{MMV\_Im2Im} is to pre-pack common bioimaging datasets as a Dataset module, so that DL researchers can use it for algorithm development and benchmarking, and new users can easily use it for learning microscopy image-to-image transformation. We will continue improving the functionalities of the package, such as supporting more models and methods, such as diffusion based models {[}\protect\hyperlink{ref-1A3yurr7m}{46}{]}, unsupervised denoising {[}\protect\hyperlink{ref-4vnyY9J9}{42}{]} or Imaginaire {[}\protect\hyperlink{ref-vzOBQiEH}{47}{]}. Besides, we also plan to develop two auxiliary packages \emph{MMV\_Im2Im\_Auto} and \emph{MMV\_Im2Im\_Active}. In specific, when you have a reasonable amount of training data, \emph{MMV\_Im2Im\_Auto} will take advantage of the fact that MMV\_Im2Im is fully configurable with yaml files, and automatically generate a set of potentially good configurations, then find the optimal solution for you by cross validation. On the other hand, when you only have very limited training data, or even with only pseudo ground truth, \emph{MMV\_Im2Im\_Active} will help to build preliminary models from the limited training data, and gradually refine the model with human-in-the-loop by active learning {[}\protect\hyperlink{ref-jM3v1UjQ}{20}{]}. All the packages will also be wrapped into napari plugins {[}\protect\hyperlink{ref-YEMgt2T4}{48}{]} to allow no-code operation and therefore be more friendly to users without experience in programming.

The image-to-image transformation frameworks implemented in the current version do not explicitly take temporal information into account. We treat images (2D or 3D) at each time step independently. Thanks to the flexibility of aicsimageio, our package can directly read even multi-channel 3D timelapse data (i.e, 5D) during training or inference, if necessary. But the computation is done at individual time steps. A common method to integrate the temporal context with spatial information is the convolutional recurrent neural network (CRNN) {[}\protect\hyperlink{ref-s2RBSHdH}{49}{]}. The support of CRNN will be part of our future work.

Another type of microscopy image analysis problem related to image-to-image transformation is image registration, where we learn how to transform the ``floating'' image spatially so that it is optimally aligned with the reference image in the physical space. Recent methods are able to transform the floating image into its registered version through deep neural networks {[}\protect\hyperlink{ref-1Fh9QLxl9}{50}{]}. This will be another important direction for future extension.

@@

\begin{itemize}
\item
  Project name: MMV\_Im2Im (Microscopy Machine Vision, Image-to-Image transformation)
\item
  Project home page: {[}\protect\hyperlink{ref-github_mmv_im2im}{\textbf{github\_mmv\_im2im?}}{]}
\item
  Operating system(s): Linux and Windows (when using GPU), also MacOS (when only using CPU)
\item
  Programming language: Python
\item
  Other requirements: PyTorch 2.0.1 or higher, PyTorch Lightning \textgreater{} 2.0.0, and all other additional dependencies are specified as in {[}\protect\hyperlink{ref-github_mmv_im2im}{\textbf{github\_mmv\_im2im?}}{]}
\item
  License: MIT license
\end{itemize}

To enhance the accessibility and traceability of our toolbox, we registered it with biotools (bio.tools ID: biotools:mmv\_im2im) and workflow hub{[}\protect\hyperlink{ref-xl7YzUeX}{51}{]}.

@@

For data source 2 {[}\protect\hyperlink{ref-8ywSgqrJ}{53}{]}, it contains two sub-folders (train and test), each with snapshots sliced from different time lapse data. Each snapshot is saved as six different tiff files (only the \_bf and the second channel of \_fluo were used in this work):

>>>>

This manuscript
(\href{https://MMV-Lab.github.io/im2im-paper/v/dd4b489eea3392962ffb0d1388911c331d929bbb/}{permalink})
was automatically generated
from \href{https://github.com/MMV-Lab/im2im-paper/tree/dd4b489eea3392962ffb0d1388911c331d929bbb}{MMV-Lab/im2im-paper@dd4b489}
on December 21, 2023.

@@

For computational biomedical researchers, \emph{MMV\_Im2Im} provides a starting point for developing new biomedical image analysis or machine learning algorithms, where they can either reuse the code in this package or fork and extend this package to facilitate the development of new methods. Experimental biomedical researchers can benefit from this work by gaining a comprehensive view of the image-to-image transformation concept through diversified examples and use cases. We hope this work can give the community inspirations on how DL-based image-to-image transformation can be integrated into the assay development process, enabling new biomedical studies that cannot be done only with traditional experimental assays. To help researchers get started, we have provided source code, documentation, and tutorials for \emph{MMV\_Im2Im} at {[}\protect\hyperlink{ref-10dtMviwb}{1}{]} under MIT license.

@@

The original idea of a general toolbox actually stemmed from the OpenMMLab project (\url{https://openmmlab.com/}), which provides generic codebases for a wide range of computer vision research topics. For instance, \emph{MMSegmentation} {[}\protect\hyperlink{ref-EcizztLg}{17}{]} is an open source toolbox for semantic segmentation, supporting unified benchmarking and state-of-the-art models ready to use out-of-box. It has become one of most widely used codebase for research in semantic segmentation (2.3K forks and 6.5K stars on GitHub as of September 29, 2023). This inspires us to develop \emph{MMV\_Im2Im} to facilitate research in image-to-image transformation with a special focus on biomedical applications.

First of all, different from general computer vision datasets, such as ImageNet {[}\protect\hyperlink{ref-lt4BNUoG}{18}{]}, where the images are usually small 2D RGB images (e.g., 3 x 256 x 256 pixels), biomedical applications usually involves large-scale high dimensional data (e.g., 500 images of 4 x 128 x 2048 x 2048 voxels). To deal with this issue, we employ the PersistentDataset in MONAI {[}\protect\hyperlink{ref-UU62HYC6}{19}{]} with partial loading and sampling support, as well as delayed image reading powered by aicsimageio {[}\protect\hyperlink{ref-gsfWGJKf}{20}{]} as default (configurable if another dataloader is preferred). As a result, in our stress test, training a 3D nuclei instance segmentation model with more than 125,000 3D images can be conducted efficiently in a day, even with limited resources.

@@

Third, the toolbox pre-packs common functionalities specific to microscopy images. For example, we incorporate the special image normalization method introduced in {[}\protect\hyperlink{ref-Yq8wZ6hc}{5}{]}, where only the middle chunk along the Z dimension of 3D microscopy images will be used for calculating the mean and standard deviation of image intensity for standard normalization. Also, 3D light microscopy images are usually anisotropic, i.e., much lower resolution along Z than XY dimension. So, we adopt the anisotropic variation of UNet as proposed in {[}\protect\hyperlink{ref-jM3v1UjQ}{21}{]}.

Finally, to deploy the model in production, a model trained on small 3D patches sometimes needs to be applied not only on much larger images. Combining the efficient data handling of aicsimageio {[}\protect\hyperlink{ref-gsfWGJKf}{20}{]} and the sliding window inference with gaussian weighted blending, the toolbox can yield efficient inference without visible stitching artifacts in production.

All in all, the \emph{MMV\_Im2Im} toolbox stands on the shoulders of many giants in the open-source software and ML engineering communities (pytorch-lightning, MONAI, aicsimageio, etc.) and is systematically designed for image-to-image transformation R\&D for biomedical applications. The source code of \emph{MMV\_Im2Im} is available at {[}\protect\hyperlink{ref-10dtMviwb}{1}{]}. This manuscript is generated with the open-source package Manubot {[}\protect\hyperlink{ref-YuJbg3zO}{22}{]}. The manuscript source code is available at {[}\protect\hyperlink{ref-B5mAMFLK}{23}{]}.

@@

We did a special comparison in this subsection to further illustrate the difference between semantic and instance segmentations. We took the 3D fibrillarin dataset from {[}\protect\hyperlink{ref-5sGcmDuy}{28}{]}. There are multiple channels in each 3D image, including DNA dye, membrane dye, and the structure channel (i.e., fibrillarin in this case). The original fibrillarin segmentation released with the dataset is a semantic segmentation (0=background, 1=fibrillarin). With the additional cell segmentation available in the dataset, we can know which groups of segmented fibrillarin belong to the same cell. Then, we can convert the original 3D fibrillarin semantic segmentation ground truth into 3D instance segmentation ground truth (fibrillarin pixels belonging to the same cell are grouped as a unique instance). Sample images and results are shown in Figure \ref{fig:3dseg}. We can observe that the semantic segmentation model is able to achieve good accuracy in determining pixels from the fibrillarin signals (F1 = 0.958 ± 0.008). Meanwhile, the instance segmentation can group them properly (average precision at 50 = 0.795 ± 0.055) so that fibrillarin masks from the same cell are successfully identified as unique instances, even without referring to the cell membrane channel or cell segmentation results. This is not a simple grouping step based on distance, since the fibrillarin signals from tightly touching nuclei may exist very close to each other.

@@

Large amounts of high-quality segmentation ground truth is not always available, or may require endless effort to collect for a segmentation task. CycleGAN-based methods have opened up a new avenue for segmentation without the need for pixel-wise ground truth {[}\protect\hyperlink{ref-RuFP3CS3}{13}{]}. In this subsection, we demonstrate an unsupervised learning-based segmentation method on four examples: 2D tight-junction (via ZO1) segmentation from 2D FP-tagged ZO1 images (max-projected from 3D stacks), and segmentation of nuclei, mitochondria, and golgi from 3D confocal microscopy images.

To perform unsupervised learning, we used raw images from the hiPSC single-cell image dataset {[}\protect\hyperlink{ref-5sGcmDuy}{28}{]}, as well as their corresponding segmentations (may not be absolute pixel-wise ground truth, but have gone through systematic evaluation to ensure the overall quality). We shuffled the raw images and their segmentations to generate a set of simulated segmentation masks. A demonstration of the concept is illustrated in Figure \ref{fig:unsupervised}-A. Example results for all 3D models are shown in Figure \ref{fig:unsupervised}-B, and the F1-scores on the test set are summarized in Table \ref{tbl:unsuper}.

For the 2D example, we saw that the unsupervised training provides a valuable segmentation, which is reflected by the F1 Score in Table \ref{tbl:unsuper}. For the 3D examples, it has been suggested that the quality of unsupervised nuclei segmentation could be further improved with additional simulation strategies {[}\protect\hyperlink{ref-RuFP3CS3}{13}{]}. Overall, we believe that unsupervised learning offers an effective way to generate preliminary segmentation, which can be further refined through active learning such as the iterative DL workflow described in {[}\protect\hyperlink{ref-jM3v1UjQ}{21}{]}.

@@

Generating a large amount of synthetic microscopy images can be an important step in developing image analysis methods. Synthetic images offer a way to train other DL models, such as self-supervised pre-training, using a diverse set of images without the need for large amounts of real-world data. As long as the synthetic images are generated with sufficient quality, it is possible to have an unlimited amount of training data for certain applications. Moreover, synthetic images can be used to evaluate other models when validation data is difficult to obtain. In this study, we demonstrate that \emph{MMV\_Im2Im} can generate 2D/3D synthetic microscopy images with high realism and validity, using a subset of data collected from the hiPSC single-cell image dataset {[}\protect\hyperlink{ref-5sGcmDuy}{28}{]}, either in a supervised or unsupervised manner.

@@

\emph{MMV\_Im2Im} can also be used to computationally reduce image noise or restore the data from various sources of imaging artifacts, so as to increase the feasibility and efficiency in downstream analysis. In the current version of \emph{MMV\_Im2Im}, the restoration model can only be trained in a fully supervised manner. Therefore, aligned low-quality and high-quality images are required for supervision, even though such paired data can be partially simulated {[}\protect\hyperlink{ref-wcCVn8av}{7}{]}. Other methods, such as unsupervised learning-based solutions {[}\protect\hyperlink{ref-4vnyY9J9}{44}{]}, will be made available within \emph{MMV\_Im2Im} in future versions.

In this example, we presented an image denoising demonstration with sample data from {[}\protect\hyperlink{ref-12G712Zky}{45}{]}. The goal was to increase the quality of low signal-to-noise ratio (SNR) images of nucleus-stained flatworms (Schmidtea mediterranea, planaria) and lightsheet images of Tribolium castaneum (red flour beetle) embryos. The models were trained with paired data acquired with low and high laser intensity on fixed samples, and then applied on live imaging data. For the nucleus-stained flatworm data (a test set of 20 images are available), the model achieved Pearson correlation of 0.392 ± 0.065, while the Pearson correlation between the noisy raw and ground truth images was 0.065 ± 0.057. For the red flour beetle dataset, the model has improved the Pearson correlation from 0.219 ± 0.045 to 0.444 ± 0.077 (6 images). Based on this and the results in Figure \ref{fig:denoising}, it can be observed that the low SNR images can be greatly improved. Systematic quantitative evaluations would be necessary to confirm the biological validity, but beyond the scope of this paper.

@@

Another important application of image-to-image transformation is imaging modality transformation {[}\protect\hyperlink{ref-UEBDZ3tI}{9}{]}, usually from one ``cheaper'' modality with lower resolution (e.g., with larger field-of-view, easier to acquire and scale up) to another modality with higher resolution but expensive to obtain. Such models will permit a new way in assay development strategy to take advantage of all the benefits of the cheaper modality with lower resolution and still be able to enhance the resolution computationally post hoc. To demonstrate the application of \emph{MMV\_Im2Im} in this scenario, we took an example dataset with paired 3D confocal and Stimulated Emission Depletion (STED) images of two different cellular structures, microtubule and nuclear pore {[}\protect\hyperlink{ref-UEBDZ3tI}{9}{]}. Sample results were summarized in Figure \ref{fig:modality_mt} and Figure \ref{fig:modality_np}. The corresponding error plots show pixel-based absolute differences between ground truth and prediction. Intensities were normalized to the interval from -1 to 1 for training, with intensity limits restricted to the 0.01-percentile and 99.99-percentile values of the intensity distribution.

@@

DL has emerged as a powerful tool for multiplex imaging, a powerful technique that enables the simultaneous detection and visualization of multiple biomolecules within a single tissue sample. This technique is increasingly being used in biomedical experiments but demands efficient image analysis solutions to accurately identify and quantify the different biomolecules of interest at scale. DL has demonstrated great potentials in analyzing multiplex datasets, as it can automatically learn the complex relationships between different biomolecules and their spatial distribution within tissues. Specifically, in this study, we present the effectiveness of \emph{MMV\_Im2Im} in transforming tissue images from one staining to another, which will permit efficient co-registration, co-localization, and quantitative analysis of multiplex datasets. We used the sample dataset from {[}\protect\hyperlink{ref-WwenuBHa}{6}{]}. In this example, we trained three different models to transform IHC images to images of standard hematoxylin stain, mpIF nuclear (DAPI) and mpIF LAP2beta (a nuclear envelope stain). Example results can be observed in Figure \ref{fig:multiplex} to verify the results qualitatively, the respective metrics can be found in Table \ref{tbl:multiplex}. It is worth mentioning that there is a pixel shift in the mpIF LAP2beta holdout dataset, but image registration is beyond the scope of this manuscript. We show the metrics as an example of an evaluation of the transformed images, but we leave an application-specific evaluation to the appropriate researchers. But it is evident that these transformed images can provide valuable insights into the localization and expression patterns of specific biomolecules spatially.

@@

From all experiments above (37 in total), we want to demonstrate the great flexibility of \emph{MMV\_Im2Im} and not to optimize every task in detail. Presenting all detailed configurations in these 37 experiments in the manuscript could lead to more confusion than clarity. To this end, we give a high-level overview of the key information of each task in Table \ref{tbl:framework_overview}, hoping to serve as a valuable starting point for researchers to optimize their DL-based image-to-image transformation using \emph{MMV\_Im2Im}. The full configuration details are available in human-readable formats in our GitHub repository {[}\protect\hyperlink{ref-10dtMviwb}{1}{]}.

@@

Overall, the package inherited the boilerplate concept from pytorch-lightning {[}\protect\hyperlink{ref-YbvSvdyB}{15}{]}, and was made fully configurable via yaml files supported by pyrallis {[}\protect\hyperlink{ref-17WhsEZko}{46}{]}, as well as largely employed state-of-the-art DL components from MONAI {[}\protect\hyperlink{ref-UU62HYC6}{19}{]}. The three key parts in the package: \texttt{mmv\_im2im.models}, \texttt{mmv\_im2im.data\_modules}, and \texttt{Trainers}, will be further described below.

@@

In general, there are mainly four major DL frameworks that could be applied to microscopy image-to-image transformation: supervised learning with a fully convolutional networks (FCN) type models, supervised learning with pix2pix type models, unsupervised learning to learn mapping between visual domains, and Self2Self-type self-supervised learning {[}\protect\hyperlink{ref-tuObtXMR}{47}{]}. The major difference between FCN based supervised learning and pix2pix based supervised learning is that the pix2pix framework extends an FCN model with an adversarial head as a discriminator to further improve the realism of the prediction. The major difference between the unsupervised framework and the self-supervised framework is that the unsupervised methods still require examples of the target images, even though the source images and target images do not need to be from the same sample or pixel-wise aligned. But, the self-supervised framework would only need the original images, which could be really helpful when it is impossible to acquire the target images (e.g., there is no truly noise-free or artifact-free image).

Currently, for supervised frameworks, both the FCN-type and pix2pix-type are well supported in the MMV\_Im2Im (RRID:SCR\_024630) package. Since our package is designed in a very generic way, it is possible to continuously expand the functionalities when available (ideally with community contributions). For example, diffusion models {[}\protect\hyperlink{ref-1A3yurr7m}{48}{]} can be thought of as a modern extension of the pix2pix-type framework and therefore are within our horizon to include into \emph{MMV\_Im2Im}. For the unsupervised framework, only CycleGAN-type methods are supported. We are planning to extend the unsupervised framework with Imaginaire {[}\protect\hyperlink{ref-vzOBQiEH}{49}{]}, which will greatly extend the applicability of \emph{MMV\_Im2Im} (e.g., learning the transformation from one single image to another single image or one set of images to another set of images). Meanwhile, supporting the self-supervised framework will be our next major milestone.

@@

One of main directions for extending \emph{MMV\_Im2Im} is to pre-pack common bioimaging datasets as a Dataset module, so that DL researchers can use it for algorithm development and benchmarking, and new users can easily use it for learning microscopy image-to-image transformation. We will continue improving the functionalities of the package, such as supporting more models and methods, such as diffusion based models {[}\protect\hyperlink{ref-1A3yurr7m}{48}{]}, unsupervised denoising {[}\protect\hyperlink{ref-4vnyY9J9}{44}{]} or Imaginaire {[}\protect\hyperlink{ref-vzOBQiEH}{49}{]}. Besides, we also plan to develop two auxiliary packages \emph{MMV\_Im2Im\_Auto} and \emph{MMV\_Im2Im\_Active}. In specific, when you have a reasonable amount of training data, \emph{MMV\_Im2Im\_Auto} will take advantage of the fact that MMV\_Im2Im is fully configurable with yaml files, and automatically generate a set of potentially good configurations, then find the optimal solution for you by cross validation. On the other hand, when you only have very limited training data, or even with only pseudo ground truth, \emph{MMV\_Im2Im\_Active} will help to build preliminary models from the limited training data, and gradually refine the model with human-in-the-loop by active learning {[}\protect\hyperlink{ref-jM3v1UjQ}{21}{]}. All the packages will also be wrapped into napari plugins {[}\protect\hyperlink{ref-YEMgt2T4}{50}{]} to allow no-code operation and therefore be more friendly to users without experience in programming.

The image-to-image transformation frameworks implemented in the current version do not explicitly take temporal information into account. We treat images (2D or 3D) at each time step independently. Thanks to the flexibility of aicsimageio, our package can directly read even multi-channel 3D timelapse data (i.e, 5D) during training or inference, if necessary. But the computation is done at individual time steps. A common method to integrate the temporal context with spatial information is the convolutional recurrent neural network (CRNN) {[}\protect\hyperlink{ref-s2RBSHdH}{51}{]}. The support of CRNN will be part of our future work.

Another type of microscopy image analysis problem related to image-to-image transformation is image registration, where we learn how to transform the ``floating'' image spatially so that it is optimally aligned with the reference image in the physical space. Recent methods are able to transform the floating image into its registered version through deep neural networks {[}\protect\hyperlink{ref-1Fh9QLxl9}{52}{]}. This will be another important direction for future extension.

@@

\begin{itemize}
\item
  Project name: MMV\_Im2Im (Microscopy Machine Vision, Image-to-Image transformation)
\item
  Project home page: {[}\protect\hyperlink{ref-10dtMviwb}{1}{]}
\item
  Operating system(s): Linux and Windows (when using GPU), also MacOS (when only using CPU)
\item
  Programming language: Python
\item
  Other requirements: PyTorch 2.0.1 or higher, PyTorch Lightning \textgreater{} 2.0.0, and all other additional dependencies are specified as in {[}\protect\hyperlink{ref-10dtMviwb}{1}{]}
\item
  License: MIT license
\end{itemize}

To enhance the accessibility and traceability of our toolbox, we registered it with biotools (bio.tools ID: biotools:mmv\_im2im) and workflow hub{[}\protect\hyperlink{ref-xl7YzUeX}{53}{]}.

@@

For data source 2 {[}\protect\hyperlink{ref-8ywSgqrJ}{55}{]}, it contains two sub-folders (train and test), each with snapshots sliced from different time lapse data. Each snapshot is saved as six different tiff files (only the \_bf and the second channel of \_fluo were used in this work):
